### 2015-03-19 text/appendix_denotational.tex
added definitions of rule and layer executions

>>>>

\subsection{Semantics of DSLTrans}

\begin{definition}{Set of Rule Executions.}
\label{def:rule_exec_set}
Let $rl\in \textsc{Rule}^{sr}_{tg}$ be a DSLTrans rule. The set of all rule
executions is a set of graph relations, built inductively as follows:
\begin{align*}
\llbracket rl\rrbracket =\big\{&(i',o')\in \textsc{Gr}\;\lvert\;(i',o') =
(rl_{\emptyset},rl_{\emptyset})\;\lor \\&(i',o') = (o,o\sqcup rl_{pat})\land
(i,o)\in \llbracket rl\rrbracket \big\}
\end{align*} 
where  $rl_{pat}$ is a pattern over $rl$ and $\sqcup$
is the tracing graph union where at least one element of $\Vert rl_{pat}\Vert$
is disjoint with $o$.
\end{definition}

A set of rule executions is the set of all input/output pairs resulting from
executing one rule over any input model. It is built inductively by starting
from $rl_{\emptyset}$ (any input graph where the rule does not match) and adding
to it any traced patterns of $rl$, meaning $rl$ has executed any number of
times. Note that the match part of $rl_{pat}$ (denoted $\Vert rl_{pat}\Vert$)
can overlap with the graph in $o$ by at most $n-1$ elements, being that $n$ is the amount
of match elements in $rl_{pat}$. Traceability adds trace links between any
elements belonging all input elements and any output elements of $rl_{pat}$ that
is not connected to a backward link.

\levi{the definition of sets of rule executions is based on the notion of
patterns over rules from Bernhard's text. I think I got the intent of
\ref{def:rule_pattern}, but more discussion is needed such that I'm sure of my
definition. In particular the notion of overlap between graphs is
underspecified regarding types (is this solved in the basic definitions?)}

\begin{definition}{Set of Rule Executions for a Layer.}
\label{def:rule_exec_set}
Let $l\in \textsc{Layer}^{sr}_{tg}$ be a DSLTrans layer. The set of all layer
executions is a set of graph relations, built inductively as follows:
\begin{gather*}
  \llbracket l\rrbracket = 
  \begin{cases}
    \emptyset &\text{if}\; l=\emptyset\\
    \llbracket l_1\rrbracket \otimes \llbracket rl\rrbracket &\text{if}\;
    l = l_1\cup rl
  \end{cases}
\end{gather*}

where $\textsc{Ex}_1 \otimes \textsc{Ex}_2$ is set of the unions of pairs of
graphs $(ex_1,ex_2)\in \textsc{Ex}_1 \times \textsc{Ex}_2$ ($\textsc{Ex}_1$ and
$\textsc{Ex}_2$ being sets of executions) where the match elements of $ex_1$ and
the match elements of $ex_2$ may be joint.
\end{definition}

\levi{this is overapproximating the set of layer executions because of the
corner case where the match of one rule contains the match of the other rule in
the layer, in which case the smaller rule will necessarily execute where the
larger one does (but not vice-versa). I think there is an elegant way to
describe this\ldots or can we just use the overapproximation as is, given it
does contain the set of all executions of the layer\ldots}
